Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems

### 2017-08-17 almo-langevin.tex
August 17th v2
Corrected a small math mistake

The concept of balancing an error in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
Previously, $\epsilon_{SCF}$ had to be set very low to ensure sufficient convergence of the electron structure calculations. 
By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
Obviously, this introduces a significant error into the calculation of forces during each time step. 
We assume that this error has the form:
%
\begin{align}
\label{eq:assumption}
f^{\text{SCF}}_{i\alpha} = f^{\text{APP}}_{i\alpha} + R^{\Delta}_{i\alpha} (t)
\end{align}
%
where $f^{\text{SCF}}_{i\alpha}$ is the fully converged ALMO DFT force calculation, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L-\Delta$.
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{APP}}_{i\alpha} - \gamma_L m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t)
\end{align}
%
Note that $\Delta$ can be either positive or negative depending on the way in which errors are introduced in the system of interest. 
$\Delta$ was found to be positive for all systems considered here, reducing the effective random force on the system.

>>>>

The concept of balancing an error in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
Previously, $\epsilon_{SCF}$ had to be set very low to ensure sufficient convergence of the electron structure calculations. 
By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
Obviously, this introduces a significant error into the calculation of forces during each time step. 
We assume that this error has the form:
%
\begin{align}
\label{eq:assumption}
f^{\text{SCF}}_{i\alpha} = f^{\text{APP}}_{i\alpha} + R^{\Delta}_{i\alpha} (t)
\end{align}
%
where $f^{\text{SCF}}_{i\alpha}$ is the fully converged ALMO DFT force calculation, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
The Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one $\gamma = \gamma_L+\Delta$.
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{APP}}_{i\alpha} - \gamma_L m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t)
\end{align}
%
Note that $\Delta$ can be either positive or negative depending on the way in which errors are introduced in the system of interest. 
$\Delta$ was found to be positive for all systems considered here, increasing the effective random force on the system.
